Fig3

### 2020-02-18 analysis/02_DESeq2.Rmd
now with prolactin hi low

```{r setup}

library(tidyverse)
library(DESeq2)

library(BiocParallel)
register(MulticoreParam(6))

@@

```{r manipulationwithin}
# import "colData" which contains sample information and "countData" which contains read counts
countData <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
head(geneinfo)

@@

>>>>

```{r setup}
library(tidyverse)
library(DESeq2)
library(BiocParallel)
register(MulticoreParam(6))
library(caret)

@@

```{r manipulationwithin, eval = F}
# import "colData" which contains sample information and "countData" which contains read counts
countData <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
head(geneinfo)

@@

## PRL-driven DESeq2

```{r PRL}
PRLdata <- read_csv("../results/PRLvsd.csv")
PRLsamples <- colData %>%
  mutate(samples = sapply(strsplit(as.character(V1), '.NYNO'), "[", 1))  %>%
  select(samples) 
PRLsamples

# join with hi lo, and drop controls
colDataPRL <- full_join(PRLdata, PRLsamples) %>% 
  drop_na() %>%
  select(samples, tissue, sex, treatment, hiloPRL) %>% 
  filter(!treatment %in% c("control", "bldg"))

# sample 100 only, better for deseq2
colDataPRL <- colDataPRL[sample(1:nrow(colDataPRL), 100,
   replace=FALSE),]

colDataPRL <- as.data.frame(colDataPRL)
row.names(colDataPRL) <- colDataPRL$samples

savecols <- as.character(colDataPRL$samples) 
savecols <- as.vector(savecols)

colDataPRL %>%
  group_by(sex, tissue, hiloPRL) %>%
  summarize(n = n())

dds <- DESeqDataSetFromMatrix(countData = newcountData,
                                colData = colDataPRL,
                                design = ~ tissue * sex * hiloPRL )
dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
print(dds)
print(dim(dds))
dds <- DESeq(dds, parallel = TRUE) # Differential expression analysis

vsd <- as.data.frame(assay(vst(dds, blind=FALSE)))
```

### 2020-02-18 analysis/figure-3.Rmd
now with prolactin hi low

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

PRLvsd <- vsd_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name")  %>% 
  dplyr::rename("gene" = "X1") %>% 
  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
               names_to = "samples", values_to = "counts") %>%
  filter(gene == "PRL")  %>%
  dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
  dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/'), "[", 2)) %>%
  dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
  dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
  dplyr::select(sex, tissue, treatment, gene, samples, counts)
PRLvsd$treatment <- factor(PRLvsd$treatment, levels = alllevels)

@@

>>>>

@@

PRLvsd <- vsd_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name")  %>% 
  dplyr::rename("gene" = "X1") %>% 
  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
               names_to = "samples", values_to = "counts") %>%
  filter(gene == "PRL")  %>%
  dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
  dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/'), "[", 2)) %>%
  dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
  dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
  dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
  drop_na()
PRLvsd$treatment <- factor(PRLvsd$treatment, levels = alllevels)

@@

```{r determinePRLhiglow}
PRLvsd2 <- PRLvsd %>%
  filter(tissue == "pituitary")
head(PRLvsd2)

PRLvsd2 %>%
  group_by(sex) %>%
  summarize(median = median(counts))

a <- ggplot(PRLvsd2, aes(x = sex, y = counts, fill = sex))  +
  geom_boxplot() +
  labs(x = "Sex", y = "Prolactin expression")+
  theme_B3() + theme(legend.position = "none") +
  scale_fill_manual(values = sexcolors) +
  geom_hline(yintercept=18, linetype="dashed", color = "black")

PRLvsd3 <- PRLvsd2 %>%
  mutate(hiloPRL = ifelse(counts >= 18, "hi", "lo"))  %>%
  drop_na()
PRLvsd3$hiloPRL <- factor(PRLvsd3$hiloPRL, levels = c("lo", "hi"))

b <- ggplot(PRLvsd3, aes(x = hiloPRL, y = counts, fill = sex))  +
  geom_boxplot() +
  labs(x = "Prolactin, binned", y = "Prolactin expression")+
  theme_B3() + theme(legend.position = "none") +
  scale_fill_manual(values = sexcolors)

plot_grid(a,b, nrow = 1, rel_widths = c(1,1))

PRLvsd3 %>%
  group_by(sex, tissue, hiloPRL) %>%
  summarize(n = n())

write.csv(PRLvsd,"../results/PRLvsd3.csv", row.names = F)
```
